Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-02-10 paper.tex

\chapter{Domain nonspecific BigData frameworks} \section{mapreduce}
\begin{itemize} \item many computations can be described through map and reduce
\item fault tolerant \end{itemize} \section{spark} \begin{itemize} \item based
on map-reduce \item in-memory processing \item lazy evaluation \item general
framework \end{itemize} \section{Use of containers with Big Data frameworks}
\section{Hadoop Distributed file system} \note{maybe should go in between
mapreduce and spark} \begin{itemize} \item enables data-locality in Spark/
decouples data locality logic from mapreduce\ \item data replication \item
fault-tolerance \end{itemize} \chapter{Scientific workflows in Spark}
\section{Introduction} \section{ariel's paper and thunders astronomy - }
\section{Spark on hpc's or } \note{large dataset for valduriez} Prior to
examining how Big Data frameworks could be improved on HPC clusters, it is
important to have a baseline of how it performs natively on an HPC cluster.
Particularly, how it behave when processing scientific workflows. In a recent
study done by \cite{valduriez}, they examined the scalability of Spark on an HPC
cluster using a black-box scientific workflow. Five tests were performed in
their study: 1) Measuring execution time when task and resources available
increase proportionally, 2) Measuring the execution time of very short and short
tasks, 3) Measuring the execution time of short (5s) to long tasks (120s), 4)
Varying the number of tasks with fixed task durations, 5) Mixed task durations
and varied number of tasks. Their results show Spark behaves as expected when
tasks and resources are increased exponentially, and that Spark scales better
than expected with tasks of longer durations. This is due to the fact that the
scheduler is overloaded in the case of many short tasks in addition to their
being a significant amount of I/O occurring between each task. Although I/O is
problematic in both data center and HPC infrastructures, HPC is optimized for
data transfer over the network rather than disk I/O, and therefore, such
frequent I/O negatively impacts the system. However, many scientific workflows
are composed of longer tasks, and therefore, Spark scales very well with
scientific workflows in an HPC environment.

@@

    Alluxio has is made up of two layers: the lineage layer and the persistence
layer. The lineage layer is the layer that performs recomputations to obtain the
desired output. The persistance layer contains all the code used to regenerate
the data, in addition to any data that cannot be recovered through lineage
as it may be too large to fit in memory or has no lineage information stored.

>>>>

\chapter{Domain nonspecific BigData frameworks} 
    \section{mapreduce}
        Summary of ~\cite{mapred}

        Typically, simple computations need to be performed on 
        increasingly growing large datasets. Processing of such 
        large datasets require parallelization. As a result, 
        these computations are complicated by details of 
        parallelization, fault tolerance, data distribution and 
        load balancing. The MapReduce library addresses this issue 
        by allowing the expression of simple computations while 
        abstracting the other details. The map and reduce paradigm 
        was selected for this purpose as it was already commonly 
        used in functional programming languages, such as Lisp, and 
        it was observed that many computations could be expressed 
        as such.

        The MapReduce library operates by first dividing the data 
        into manageable chunks (16 to 64MB). It then creates 
        multiple copies of the program and distributes them across
        nodes. There exists a single master node; the remainder are
        worker nodes. The master node is responsible for delegating 
        map and reduce tasks to the idle workers. There are two data
        structures found in the master. The first stores the state 
        (\textit{idle}, \textit{in-progress} or \textit{completed}) 
        of each \textit{map} and \textit{reduce} task, and the second
        stores the identity of the non-idle worker nodes. In contrast,
        the worker nodes are responsible for the execution of 
        \textit{map} or \textit{reduce} tasks. A worker node assigned
        a \textit{map} task will parse out the key-value pairs from 
        the input file and pass it to the \textit{map} function. The 
        intermediate keys produced by the \textit{map} function are 
        buffered in memory and periodically written to disk.The 
        locations of these intermediate files are sent to the master 
        who is responsible for forwarding these locations to the 
        reducer. A worker node assigned a \textit{reduce} task uses 
        remote procedure calls to obtain the intermediate local files
        created by the map tasks and sorts the keys such that identical
        keys are grouped together. Each unique key and its list of 
        values are then submitted to the \textit{reduce} task and the 
        output of the \textit{reduce} task is stored in a final output 
        file. Once all \textit{reduce} tasks have completed, the master
        resumes the program and the output is returned to the user.

        In order to ensure fault tolerance, the master regularly pings 
        the worker nodes. Should a worker node not respond after a certain
        delay, the master labels the worker as failed. The \textit{map} 
        tasks that were completed by the worker are all returned to their
        original state (\textit{idle}) as \textit{map} task data is stored
        locally. Both the completed and in-progress worker's \textit{map}
        tasks are subsequently reassigned to other workers. Only the failed
        worker's \textit{in-progress} \textit{reduce} tasks need to be 
        reassigned.

        The master node undergoes periodic checkpoints, such that the 
        master node can be reinitialized from its last checkpoint in 
        case of failure. Master node failure is, however, unlikely.

        As network bandwidth is a scarce resource, the master attempts 
        to schedule \textit{map} tasks closest to where the data is 
        located. it will first attempt to schedule a map task on the host
        that contains the data; should that option not be available, the
        master will attempt to schedule the \textit{map} task on a host 
        nearest to the data (e.g. same network switch)

        Ideally, the number of \textit{map} and \textit{reduce} tasks 
        should be much larger than the number of available worker nodes. 
        Having it as such improves both dynamic load balancing and recovery 
        time. The number of \textit{map} tasks used in practice typically 
        corresponds to the size of the input chunks, whereas the number of 
        \textit{reduce} tasks are a small multiple of the number of worker 
        nodes used.

        When nearing completion of the MapReduce program, it is possible for
        ``stragglers" (\textit{in-progress} tasks running on nodes with below
        average performance) to delay completion. As a measure to counteract 
        this, the master node assigns the same task to idle workers. The task 
        is then marked as completed as soon as either the primary or the 
        backups complete the task. This strategy has been found to 
        significantly improve performance when used with very large datasets. 
        \note{44\% speedup with their sort program}.

        A few extensions have also been added to the library in order to 
        improve user experience. An important extension is the introduction 
        of a \textit{combiner}, which enables the commutative and associative 
        reduce function to be applied to the same keys on the hosts where the 
        \textit{map} task was executed. Like the \textit{map} task, the 
        \textit{combiner} produces intermediate output which is sent to the 
        \textit{reducer}. This alleviates the amount of key-value pairs that 
        must be transferred over the network from the map host to the reduce 
        host, and as a result, improves performance. Other extensions include 
        the ability to create user-defined partitioning functions, a guaranteed 
        increasing key sort order, the ability for the user to specify input 
        types, an option to skip records that continuously result in execution 
        failure, the ability to produce auxiliary output files in \textit{map} 
        or \textit{reduce} task, live status updates stream to HTTP server,
        the ability to run all code on a single local machine, and a 
        \textit{Counter} class.
    \section{spark} 
        \begin{itemize} 
            \item based on map-reduce 
            \item in-memory processing 
            \item lazy evaluation 
            \item general framework 
        \end{itemize} 
     \section{Use of containers with Big Data frameworks}
     \section{Hadoop Distributed file system} 
        \note{maybe should go in between mapreduce and spark}

        The Hadoop Distributed File System (HDFS)\cite{hadoop} 
        is the filesystem component of Hadoop. The metadata in 
        HDFS is stored in on a dedicated server known as the 
        NameNode, whereas application data is distributed across 
        many servers referred to as DataNode. To ensure fault 
        tolerance, HDFS replicates the data across multiple 
        DataNodes (default replication factor: 3). Not only does 
        this ensure that the data is not lost in the event of node
        failure, but it also increases data transfer bandwidth, as
        the data can be accessed from multiple nodes, and thus, 
        there are more opportunities for computations to be 
        performed nearest to where the data is located.

        The NameNode contains the namespace tree, a hierarchy of 
        directories and their files, as well as the mapping of the 
        split file blocks to DataNodes. The entire namespace is 
        stored in memory. Information on the directories and files,
        such as modification and access times, namespace and disk 
        quotes, are stored within \textit{inodes} on the NameNode. 
        The name system's metadata, \textit{inode} and file block 
        mapping, is known as the \textit{image}. Persistent record 
        of the image that is stored on the local file system are 
        known as \textit{checkpoint}. Locations of block replicas 
        are not stored in the checkpoint as they may change over 
        time. The \textit{journal} is a log of the modifications 
        made to the image. Both the checkpoint and the journal may 
        be copied across servers for increased durability. The 
        journal is played back during the NameNode restart in order 
        to restore the cluster.

        When a client requests to read data, it must first contact 
        the NameNode. The NameNode provides it with the replica 
        locations, and the client reads from the DataNode located 
        closest to it. When the client requests to write a file, it
        contacts the NameNode which selects the DataNodes that will 
        host the replicas. The client subsequently writes the data 
        directly to the DataNodes in a pipeline fashion (The data 
        gets propagated through each DataNode by being transferred
        from the nearest DataNode to it).

        There is only one NameNode assigned to each cluster, however, 
        each cluster can have multiple clients and execute multiple 
        tasks concurrently.

        A block replica stored on a DataNode consists of two files 
        stored on the local file system: the metadata and the data.

        The DataNode connects to the NameNode during startup and 
        performs handshake to for DataNode namespace ID and 
        software version verification. Should either namespace ID 
        or software version not match those of the NameNode, the 
        DataNode shuts down.

        To ensure integrity of the system, a \textit{namespace ID} 
        is assigned to all nodes during formatting of the namespace. 
        DataNodes with different namespace IDs to the NameNode will 
        not be permitted to join the cluster. Should the DataNode
        have not been assigned a namespace ID, it will be permitted 
        to join the cluster. The cluster's namespace ID will then be 
        assigned to that DataNode.

        Following the handshake, the DataNode register with the 
        NameNode using their storage ID. The storage ID is a unique 
        ID generated after initial registration to the NameNode and 
        that is persistently stored on the DataNode. It permits
        identification of the DataNode in the event of an IP address 
        or port change.

\chapter{Scientific workflows in Spark}
    \section{Introduction} 
    \section{ariel's paper and thunders astronomy - }
    \section{Spark on hpc's or } 
        \note{large dataset for valduriez} Prior to
examining how Big Data frameworks could be improved on HPC clusters, it is
important to have a baseline of how it performs natively on an HPC cluster.
Particularly, how it behave when processing scientific workflows. In a recent
study done by \cite{valduriez}, they examined the scalability of Spark on an HPC
cluster using a black-box scientific workflow. Five tests were performed in
their study: 1) Measuring execution time when task and resources available
increase proportionally, 2) Measuring the execution time of very short and short
tasks, 3) Measuring the execution time of short (5s) to long tasks (120s), 4)
Varying the number of tasks with fixed task durations, 5) Mixed task durations
and varied number of tasks. Their results show Spark behaves as expected when
tasks and resources are increased exponentially, and that Spark scales better
than expected with tasks of longer durations. This is due to the fact that the
scheduler is overloaded in the case of many short tasks in addition to their
being a significant amount of I/O occurring between each task. Although I/O is
problematic in both data center and HPC infrastructures, HPC is optimized for
data transfer over the network rather than disk I/O, and therefore, such
frequent I/O negatively impacts the system. However, many scientific workflows
are composed of longer tasks, and therefore, Spark scales very well with
scientific workflows in an HPC environment.

@@

    Alluxio has is made up of two layers: the lineage layer and the persistence
layer. The lineage layer is the layer that performs recomputations to obtain the
desired output. The persistence layer contains all the code used to regenerate
the data, in addition to any data that cannot be recovered through lineage
as it may be too large to fit in memory or has no lineage information stored.
